---
title: "SYNOP Messages"
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

## Overview

SYNOP is a standardized format for reporting weather observations
usually every hour. This vignette demonstrates how to work with SYNOP
FM-12 meteorological messages using the **climate** package.

The package provides two complementary functions for handling SYNOP
data:

- **`synop_parser()`** ŌĆö parse raw SYNOP message strings into structured
  data
- **`meteo_ogimet(source = "synop")`** ŌĆö download raw SYNOP messages
  from Ogimet web service and decode them automatically

## The `synop_parser()` Function

The `synop_parser()` function decodes raw SYNOP messages into either a
tidy data frame or a nested list structure.

### Basic Usage - data frame output:

```{r, eval=TRUE, include=FALSE}
suppressMessages(library(climate))
suppressMessages(library(knitr))

```

SYNOP messages should be provided as text vector of any length, e.g:

```{r simple_synop_multi, eval=TRUE, echo=TRUE, include=TRUE, paged.print=TRUE}
# Parse multiple messages
messages = c(
  "AAXX 01004 88889 12782 61506 10094 20047 30111 40197 53007 60001 81541",
  "AAXX 01006 12330 32782 61506 10094 20047 30111 40197 53007 60001 81541"
)

# Multiple messages as a data frame
df_multiple = synop_parser(messages, as_data_frame = TRUE)
df_multiple$source = NULL # remove raw message from output
# Display the results in a table
kable(t(df_multiple), caption = "Examplary results of parsing multiple SYNOP message into data frame")
```

### Available Columns

- When `as_data_frame = TRUE`, the returned data frame includes the
  following columns:

  - **Identification and metadata:** - `station_type` ŌĆö station
    classification code - `station_id` ŌĆö WMO station identifier -
    `region` ŌĆö geographic region code - `obs_day` ŌĆö observation day of
    month - `obs_hour` ŌĆö observation hour (UTC)

  - **Wind:** - `wind_unit` ŌĆö either "m/s" or "knots" - `wind_estimated`
    ŌĆö whether wind is estimated - `wind_direction` ŌĆö direction in
    degrees (0ŌĆō360┬░) - `wind_speed` ŌĆö speed in the reported unit

  - **Temperature and moisture:** - `air_temperature` ŌĆö air temperature
    (┬░C) - `dewpoint_temperature` ŌĆö dew point (┬░C)

  - **Pressure:** - `station_pressure` ŌĆö station-level pressure (hPa) -
    `sea_level_pressure` ŌĆö sea-level pressure (hPa) -
    `pressure_tendency` ŌĆö pressure change indicator - `pressure_change`
    ŌĆö 3-hour pressure change (hPa)

  - **Clouds:** - `cloud_cover` ŌĆö total cloud cover (oktas, 0ŌĆō8) -
    `low_cloud_type` ŌĆö type of low clouds (Sc, St, Cu, Cb) -
    `middle_cloud_type` ŌĆö type of middle clouds (Ac, As, Ns) -
    `high_cloud_type` ŌĆö type of high clouds (Ci, Cc, Cs) -
    `low_cloud_amount` ŌĆö cover of low clouds (oktas) - `visibility` ŌĆö
    horizontal visibility (m) - `cloud_base_min`, `cloud_base_max` ŌĆö
    cloud base height range (m)

  - **Precipitation and other:** - `precipitation_amount` ŌĆö total
    precipitation (mm) - `precipitation_time` ŌĆö precipitation period
    (hours) - `maximum_temperature` ŌĆö daily maximum (┬░C, 
    from Section 3) - `minimum_temperature` ŌĆö daily minimum (┬░C, from Section 3) -
    `gust` ŌĆö highest wind gust (in the message's wind unit) -
    `sunshine_duration` ŌĆö daily sunshine duration (hours) - `snow_depth`
    ŌĆö total snow depth (cm) - `snow_depth_state` ŌĆö descriptive
    snow/ground state

### Nested List Output

For more granular access most of decoded field, use the default list
output:

```{r echo=TRUE, include = TRUE, eval = TRUE}
# Single message as a nested list
synop_code = messages[1]
decoded = synop_parser()
str(decoded)
```

To parse multiple messages and keep each as a separate list element:

```{r}
# Returns a list of two decoded messages
decoded_list = synop_parser(messages, simplify = FALSE)

# Access the first message
decoded_list[[1]]
```

### Country-Specific Decoding

Some SYNOP groups (e.g., precipitation indicators) are country-specific.
Pass the `country` parameter to adjust decoding:

```{r echo=TRUE}
# Parse with Russia-specific precipitation decoding
decoded_ru = synop_parser(synop_code, country = "RU", as_data_frame = TRUE)

# Compare with default
decoded_default = synop_parser(synop_code, as_data_frame = TRUE)
```

## The `meteo_ogimet()` Function ŌĆö SYNOP Backend

The `meteo_ogimet()` function (with `source = "synop"`) downloads SYNOP
messages from [Ogimet](http://www.ogimet.com) and automatically decodes
them using `parser()`.

### Station Mode: Single or Multiple Stations

Fetch SYNOP data from OGIMET service for one or more WMO station IDs:

```{r echo=TRUE, include = TRUE, eval = TRUE}
# can be used with multiple stations:
two_stations = meteo_ogimet(
  source  = "synop",
  station = c(12330, 12375),
  date    = c("2024-01-01", "2024-01-10")
)

kable(head(two_stations))
```

### Country mode: all stations in a country

Fetch SYNOP data for all Ogimet stations in a country for a given time
range:

```{r, eval=TRUE, include=TRUE, include = TRUE}
# All Polish stations for a single day
poland = meteo_ogimet(
  source       = "synop",
  country_name = "Poland",
  date         = c("2024-01-15", "2024-01-15")
)
head(poland)
```

### Simplified vs. full data frame output

By default, `meteo_ogimet()` returns a simplified data frame with the
most commonly used variables. To get all decoded SYNOP fields, use
`simplified = FALSE`:

```{r echo=TRUE, eval = TRUE, include = TRUE}
# Full synop_parser output
full = meteo_ogimet(
  source     = "synop",
  station    = 12330,
  date       = c("2024-01-01", "2024-01-05"),
  simplified = FALSE
)
head(full)
```

To obtain both the simplified and full data frames in one call, use
`return_list = TRUE`:

```{r echo=TRUE}
result = meteo_ogimet(
  source      = "synop",
  station     = 12330,
  date        = c("2024-01-01", "2024-01-05"),
  return_list = TRUE
)
# Compact 20-column view
names(result$data)
# Full 30+ column parser output
names(result$full)
```

## Notes and Limitations

- **Ogimet server limits:** The Ogimet server caps responses at 200,000
  rows. If a date range would exceed this, the function automatically
  splits the range and recursively fetches each half.
- **Station delays:** When fetching multiple stations in station mode, a
  20-second delay is inserted between requests to avoid overloading the
  Ogimet server.
- **Missing observations:** Not all SYNOP variables are reported at all
  times. Missing fields are returned as `NA`.
- **Historical data:** Ogimet typically provides historical SYNOP data
  since approx. 1999/2000; availability varies by station and country.

## See Also

- Daily and hourly OGIMET data via the HTML backend:
  `meteo_ogimet(source = "html")`
- WMO SYNOP documentation for FM-12: [FM 12-XIII Ext.
  SYNOP](https://library.wmo.int/records/item/35713-manual-on-codes-volume-i-1-international-codes)
