---
title: "SYNOP Messages - parsing and/or downloading raw messages"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

## Overview

This vignette demonstrates how to work with SYNOP meteorological messages using the
**climate** package. SYNOP is a standardized format for reporting weather observations usually every hour. The package provides two complementary functions:

- **`parser()`** — parse raw SYNOP message strings into structured data
- **`meteo_ogimet_synop()`** — download SYNOP messages from Ogimet and decode them automatically

## The `parser()` Function

The `parser()` function decodes raw SYNOP messages into either a nested list structure or a tidy data frame.

### Basic Usage

```{r}
library(climate)

# A simple SYNOP message
synop_code = "AAXX 01004 88889 12782 61506 10094 20047 30111 40197 53007 60001 81541"

# Parse into a nested list
decoded = parser(synop_code)
str(decoded, max.level = 2)
```

### Multiple Messages

To parse multiple SYNOP messages at once:

```{r}
# Parse multiple messages
messages = c(
  "AAXX 01004 88889 12782 61506 10094 20047 30111 40197 53007 60001 81541",
  "AAXX 01006 12330 32782 61506 10094 20047 30111 40197 53007 60001 81541"
)

# Returns a list of two decoded messages
decoded_list = parser(messages, simplify = FALSE)

# Access the first message
decoded_list[[1]]
```

### Data Frame Output

For analysis and visualization, use `as_data_frame = TRUE` to get a tidy data frame:

```{r}
# Single message as a data frame
df_single = parser(synop_code, as_data_frame = TRUE)
head(df_single)

# Multiple messages as a data frame
df_multiple = parser(messages, as_data_frame = TRUE)
head(df_multiple)
```

### Available Columns

When `as_data_frame = TRUE`, the returned data frame includes the following columns:

**Identification and metadata:**
- `station_type` — station classification code
- `station_id` — WMO station identifier
- `region` — geographic region code
- `obs_day` — observation day of month
- `obs_hour` — observation hour (UTC)

**Wind:**
- `wind_unit` — either "m/s" or "knots"
- `wind_estimated` — whether wind is estimated
- `wind_direction` — direction in degrees (0–360°)
- `wind_speed` — speed in the reported unit

**Temperature and moisture:**
- `air_temperature` — air temperature (°C)
- `dewpoint_temperature` — dew point (°C)

**Pressure:**
- `station_pressure` — station-level pressure (hPa)
- `sea_level_pressure` — sea-level pressure (hPa)
- `pressure_tendency` — pressure change indicator
- `pressure_change` — 3-hour pressure change (hPa)

**Clouds:**
- `cloud_cover` — total cloud cover (oktas, 0–8)
- `low_cloud_type` — type of low clouds (Sc, St, Cu, Cb)
- `middle_cloud_type` — type of middle clouds (Ac, As, Ns)
- `high_cloud_type` — type of high clouds (Ci, Cc, Cs)
- `low_cloud_amount` — cover of low clouds (oktas)
- `visibility` — horizontal visibility (m)
- `cloud_base_min`, `cloud_base_max` — cloud base height range (m)

**Precipitation and other:**
- `precipitation_amount` — total precipitation (mm)
- `precipitation_time` — precipitation period (hours)
- `maximum_temperature` — daily maximum (°C, from Section 3)
- `minimum_temperature` — daily minimum (°C, from Section 3)
- `gust` — highest wind gust (in the message's wind unit)
- `sunshine_duration` — daily sunshine duration (hours)
- `snow_depth` — total snow depth (cm)
- `snow_depth_state` — descriptive snow/ground state

### Country-Specific Decoding

Some SYNOP groups (e.g., precipitation indicators) are country-specific. Pass the `country` parameter to adjust decoding:

```{r}
# Parse with Russia-specific precipitation decoding
decoded_ru = parser(synop_code, country = "RU", as_data_frame = TRUE)

# Compare with default
decoded_default = parser(synop_code, as_data_frame = TRUE)
```

## The `meteo_ogimet_synop()` Function

The `meteo_ogimet_synop()` function downloads SYNOP messages from [Ogimet](http://www.ogimet.com) and automatically decodes them using `parser()`.

### Station Mode: Single or Multiple Stations

Fetch SYNOP data for one or more WMO station IDs:

```{r}
# Single station: Poznan-Lawica (Poland), WMO ID 12330
poznan = meteo_ogimet_synop(
  station = 12330,
  date = c("2024-01-01", "2024-01-10")
)
head(poznan)

# Multiple stations
two_stations = meteo_ogimet_synop(
  station = c(12330, 12375),
  date = c("2024-01-01", "2024-01-10")
)
head(two_stations)
```

### Country Mode: All Stations in a Country

Fetch SYNOP data for all Ogimet stations in a country:

```{r}
# All Polish stations for a single day
poland = meteo_ogimet_synop(
  country_name = "Poland",
  date = c("2024-01-15", "2024-01-15")
)
head(poland)

# Typically returns fewer rows than expected due to missing observations
```

### Simplified vs. Full Output

By default, `meteo_ogimet_synop()` returns a simplified 20-column data frame with the most commonly used variables:

```{r}
# Simplified output (default)
simplified = meteo_ogimet_synop(
  station = 12330,
  date = c("2024-01-01", "2024-01-05"),
  simplified = TRUE
)
colnames(simplified)
```

To get all decoded SYNOP fields, use `simplified = FALSE`:

```{r}
# Full parser output
full = meteo_ogimet_synop(
  station = 12330,
  date = c("2024-01-01", "2024-01-05"),
  simplified = FALSE
)
colnames(full)
```

### Handling Network Failures

By default, network errors are caught and a message is printed instead of stopping:

```{r}
# Network-robust by default
result = meteo_ogimet_synop(
  station = 12330,
  date = c("2024-01-01", "2024-01-05"),
  allow_failure = TRUE  # default
)

# To see full error details, set allow_failure = FALSE
result = meteo_ogimet_synop(
  station = 12330,
  date = c("2024-01-01", "2024-01-05"),
  allow_failure = FALSE
)
```

## Practical Example: Temperature and Humidity Analysis

Combine simplified output with standard data manipulation:

```{r}
library(dplyr)

# Download SYNOP data for a week
data = meteo_ogimet_synop(
  station = 12330,
  date = c("2024-01-01", "2024-01-07"),
  simplified = TRUE
)

# Basic statistics
data %>%
  summarise(
    mean_temp = mean(t2m, na.rm = TRUE),
    max_temp = max(t2m, na.rm = TRUE),
    min_temp = min(t2m, na.rm = TRUE),
    mean_rh = mean(rel_hum, na.rm = TRUE)
  )

# Track pressure changes
data %>%
  arrange(date) %>%
  mutate(slp_change = slp - lag(slp)) %>%
  select(date, slp, slp_change) %>%
  head()
```

## Country-Specific Precipitation

When downloading from a specific country, use the `country` parameter for accurate precipitation decoding:

```{r}
# Russian stations with Russia-specific precipitation decoding
russia = meteo_ogimet_synop(
  country_name = "Russia",
  country = "RU",
  date = c("2024-01-15", "2024-01-15"),
  simplified = FALSE
)
```

## Notes and Limitations

- **Ogimet server limits:** The Ogimet server caps responses at 200,000 rows. If a date range would exceed this, the function automatically splits the range and recursively fetches each half.
- **Station delays:** When fetching multiple stations in station mode, a 20-second delay is inserted between requests to avoid overloading the Ogimet server.
- **Missing observations:** Not all SYNOP variables are reported at all times. Missing fields are returned as `NA`.
- **Historical data:** Ogimet typically provides several years of historical SYNOP data; availability varies by station and country.

## See Also

- Older OGIMET data (hourly/daily reanalysis): `meteo_ogimet()`
- WMO SYNOP documentation: [FM 12-XIII Ext. SYNOP](https://library.wmo.int/doc_num.php?explnum_id=3457)
