---
title: "SYNOP Messages"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

## Overview

SYNOP is a standardized format for reporting weather observations usually every hour. 
This vignette demonstrates how to work with SYNOP FM-12 meteorological messages using the
**climate** package. 

The package provides two complementary functions for handling SYNOP data:

- **`parser()`** — parse raw SYNOP message strings into structured data
- **`meteo_ogimet_synop()`** — download raw SYNOP messages from Ogimet and decode them automatically

## The `parser()` Function

The `parser()` function decodes raw SYNOP messages into either a tidy data frame or a nested list structure.

### Basic Usage - data frame output:


```{r}
suppressMessages(library(climate))
```


```{r echo=TRUE}
# A simple SYNOP message
synop_code = "AAXX 01004 88889 12782 61506 10094 20047 30111 40197 53007 60001 81541"

# Parse into a data frame (default for analysis)
df_single = parser(synop_code, as_data_frame = TRUE)
head(df_single)
```

### Multiple Messages

To parse multiple SYNOP messages at once:

```{r echo=TRUE}
# Parse multiple messages
messages = c(
  "AAXX 01004 88889 12782 61506 10094 20047 30111 40197 53007 60001 81541",
  "AAXX 01006 12330 32782 61506 10094 20047 30111 40197 53007 60001 81541"
)

# Multiple messages as a data frame
df_multiple = parser(messages, as_data_frame = TRUE)
head(df_multiple)
```

### Available Columns

When `as_data_frame = TRUE`, the returned data frame includes the following columns:

**Identification and metadata:**
- `station_type` — station classification code
- `station_id` — WMO station identifier
- `region` — geographic region code
- `obs_day` — observation day of month
- `obs_hour` — observation hour (UTC)

**Wind:**
- `wind_unit` — either "m/s" or "knots"
- `wind_estimated` — whether wind is estimated
- `wind_direction` — direction in degrees (0–360°)
- `wind_speed` — speed in the reported unit

**Temperature and moisture:**
- `air_temperature` — air temperature (°C)
- `dewpoint_temperature` — dew point (°C)

**Pressure:**
- `station_pressure` — station-level pressure (hPa)
- `sea_level_pressure` — sea-level pressure (hPa)
- `pressure_tendency` — pressure change indicator
- `pressure_change` — 3-hour pressure change (hPa)

**Clouds:**
- `cloud_cover` — total cloud cover (oktas, 0–8)
- `low_cloud_type` — type of low clouds (Sc, St, Cu, Cb)
- `middle_cloud_type` — type of middle clouds (Ac, As, Ns)
- `high_cloud_type` — type of high clouds (Ci, Cc, Cs)
- `low_cloud_amount` — cover of low clouds (oktas)
- `visibility` — horizontal visibility (m)
- `cloud_base_min`, `cloud_base_max` — cloud base height range (m)

**Precipitation and other:**
- `precipitation_amount` — total precipitation (mm)
- `precipitation_time` — precipitation period (hours)
- `maximum_temperature` — daily maximum (°C, from Section 3)
- `minimum_temperature` — daily minimum (°C, from Section 3)
- `gust` — highest wind gust (in the message's wind unit)
- `sunshine_duration` — daily sunshine duration (hours)
- `snow_depth` — total snow depth (cm)
- `snow_depth_state` — descriptive snow/ground state

### Nested List Output

For more granular access most of decoded field, use the default list output:

```{r echo=TRUE}
# Single message as a nested list
decoded = parser(synop_code)
str(decoded)
```

To parse multiple messages and keep each as a separate list element:

```{r}
# Returns a list of two decoded messages
decoded_list = parser(messages, simplify = FALSE)

# Access the first message
decoded_list[[1]]
```

### Country-Specific Decoding

Some SYNOP groups (e.g., precipitation indicators) are country-specific. Pass the `country` parameter to adjust decoding:

```{r echo=TRUE}
# Parse with Russia-specific precipitation decoding
decoded_ru = parser(synop_code, country = "RU", as_data_frame = TRUE)

# Compare with default
decoded_default = parser(synop_code, as_data_frame = TRUE)
```

## The `meteo_ogimet_synop()` Function

The `meteo_ogimet_synop()` function downloads SYNOP messages from [Ogimet](http://www.ogimet.com) and automatically decodes them using `parser()`.

### Station Mode: Single or Multiple Stations

Fetch SYNOP data for one or more WMO station IDs:

```{r echo=TRUE}
# can be used with multiple stations:
two_stations = meteo_ogimet_synop(
  station = c(12330, 12375),
  date = c("2024-01-01", "2024-01-10")
)
head(two_stations)
```

### Country mode: all stations in a country

Fetch SYNOP data for all Ogimet stations in a country for a given time range:

```{r}
# All Polish stations for a single day
poland = meteo_ogimet_synop(
  country_name = "Poland",
  date = c("2024-01-15", "2024-01-15")
)
head(poland)
```

### Simplified vs. full data frame output

By default, `meteo_ogimet_synop()` returns a simplified several-column data frame with the most commonly used variables. To get more detailed info from decoded SYNOP fields, use `simplified = FALSE`:

```{r echo=TRUE}
# Full parser output
full = meteo_ogimet_synop(
  station = 12330,
  date = c("2024-01-01", "2024-01-05"),
  simplified = FALSE
)
head(full)
```



## Notes and Limitations

- **Ogimet server limits:** The Ogimet server caps responses at 200,000 rows. If a date range would exceed this, the function automatically splits the range and recursively fetches each half.
- **Station delays:** When fetching multiple stations in station mode, a 20-second delay is inserted between requests to avoid overloading the Ogimet server.
- **Missing observations:** Not all SYNOP variables are reported at all times. Missing fields are returned as `NA`.
- **Historical data:** Ogimet typically provides historical SYNOP data since approx. 1999/2000; availability varies by station and country.

## See Also

- Older OGIMET data (hourly/daily) downloaded from the tabular form presented in the OGIMET service: `meteo_ogimet()`
- WMO SYNOP documentation for FM-12: [FM 12-XIII Ext. SYNOP](https://library.wmo.int/records/item/35713-manual-on-codes-volume-i-1-international-codes)
