---
title: "Use Case - Ogimet database"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

The `meteo_ogimet()` function is the unified entry point for downloading data from [Ogimet](https://www.ogimet.com/).
It uses two backends automatically selected by the `interval` argument:

- **`interval = "hourly"`** → **SYNOP backend** (default): downloads and decodes raw SYNOP messages via
  the Ogimet `getsynop` endpoint. Returns clean SI-unit columns (`ws` in m/s, `wd` in degrees, `t2m`
  in °C, POSIXct `date` in UTC).
- **`interval = "daily"`** → **HTML backend** (default): scrapes pre-formatted daily summary tables
  from the Ogimet `gsynres` endpoint.

Use `source = "synop"` or `source = "html"` to override the default for any interval.

---

## 1. Hourly wind patterns over Svalbard (SYNOP backend)

Download a full year of hourly observations and visualize wind patterns and temperature advection
using the [openair](https://davidcarslaw.github.io/openair/) package.

With the SYNOP backend the output already contains:

- `ws` — wind speed in **m/s**
- `wd` — wind direction in **degrees** (no character-to-degrees conversion needed)
- `t2m` — air temperature in °C
- `date` — POSIXct timestamp in UTC (the column name openair expects)

```{r svalbard-download}
library(climate)

# Download hourly data for Svalbard Lufthavn (WMO 01008)
df = meteo_ogimet(interval = "hourly",
                  date     = c("2018-01-01", "2018-12-31"),
                  station  = "01008")
head(df[, c("date", "station", "t2m", "ws", "wd")])
```

```{r svalbard-windrose, message=FALSE, warning=FALSE}
library(openair)

# Wind rose by season — no unit conversion needed
windRose(mydata  = df,
         ws      = "ws",
         wd      = "wd",
         type    = "season",
         paddle  = FALSE,
         main    = "Svalbard Lufthavn (2018)",
         ws.int  = 3,
         dig.lab = 3,
         layout  = c(4, 1))
```

```{r svalbard-timeplot}
# Temporal overview of temperature and wind speed
timePlot(df, pollutant = c("t2m", "ws"))
```

```{r svalbard-polar}
# Which wind sectors bring warm / cold air masses?
polarPlot(df,
          pollutant     = "t2m",
          x             = "ws",
          wd            = "wd",
          k             = 50,
          force.positive = FALSE,
          type          = "season",
          layout        = c(4, 1),
          resolution    = "fine",
          normalise     = FALSE)
```

---

## 2. Daily summaries — multi-station temperature comparison

The HTML backend (`interval = "daily"`) returns pre-aggregated daily statistics (Tmax, Tmin, Tavg,
precipitation, etc.) for one or more stations.

```{r daily-download}
library(climate)

# Daily summaries for two Polish stations: Poznan (12330) and Warsaw (12375)
daily = meteo_ogimet(interval = "daily",
                     date     = c("2023-06-01", "2023-08-31"),
                     station  = c(12330, 12375),
                     coords   = TRUE)
head(daily)
```

```{r daily-plot, message=FALSE}
library(ggplot2)

# Compare average temperatures between the two stations
ggplot(daily, aes(x = as.Date(Date), y = TemperatureCAvg,
                  colour = factor(station_ID), group = station_ID)) +
  geom_line() +
  labs(title  = "Daily average temperature — summer 2023",
       x      = "Date",
       y      = "Temperature (°C)",
       colour = "Station (WMO ID)") +
  theme_bw()
```

---

## 3. Country-level bulk download

The SYNOP backend supports downloading all Ogimet stations for an entire country in a single
request via the `country_name` argument. This is useful for spatial analysis.

```{r country-download}
library(climate)

# All stations in Poland for a single day
poland = meteo_ogimet(interval      = "hourly",
                      country_name  = "Poland",
                      date          = c("2023-12-15", "2023-12-15"))

cat("Stations:", length(unique(poland$station)), "\n")
cat("Observations:", nrow(poland), "\n")
head(poland[, c("date", "station", "t2m", "ws", "slp")])
```

```{r country-map, message=FALSE, warning=FALSE}
# Quick spatial overview — requires coordinates from stations_ogimet()
library(dplyr)

station_meta = stations_ogimet(country = "Poland", date = Sys.Date())

# Join decoded observations with coordinates
poland_geo = poland %>%
  group_by(station) %>%
  summarise(t2m_mean = mean(t2m, na.rm = TRUE), .groups = "drop") %>%
  left_join(station_meta, by = c("station" = "wmo_id"))

ggplot(poland_geo, aes(x = lon, y = lat, colour = t2m_mean, size = t2m_mean)) +
  geom_point(alpha = 0.7) +
  scale_colour_distiller(palette = "RdYlBu", direction = -1, name = "T2m (°C)") +
  coord_quickmap() +
  labs(title = "Mean temperature — Poland, 2023-12-15") +
  theme_bw() +
  theme(legend.position = "right")
```

---

## 4. Full SYNOP output with `return_list`

Setting `return_list = TRUE` returns a named list with two data frames: `$data` (the compact
20-column simplified view) and `$full` (the complete parser output with 30+ columns including
cloud types, precipitation period, snow state, etc.).

```{r return-list}
library(climate)

result = meteo_ogimet(interval     = "hourly",
                      station      = 12330,
                      date         = c("2023-06-01", "2023-06-03"),
                      return_list  = TRUE)

# Compact view
names(result$data)

# Full parser output
names(result$full)

# Variables only available in the full output
result$full[1, c("low_cloud_type", "middle_cloud_type", "high_cloud_type",
                  "precipitation_time", "snow_depth_state", "source")]
```

---

## 5. Selecting the backend explicitly

Override the automatic backend selection with the `source` argument:

```{r backend-override}
library(climate)

# Force SYNOP backend for daily data
poznan_synop = meteo_ogimet(interval = "daily",
                             station  = 12330,
                             date     = c("2023-06-01", "2023-06-07"),
                             source   = "synop")
head(poznan_synop[, c("date", "station", "t2m", "tmax", "tmin", "precip")])

# Force HTML backend for hourly data (returns km/h wind speed, character wind direction, etc.)
poznan_html = meteo_ogimet(interval = "hourly",
                            station  = 12330,
                            date     = c("2023-06-01", "2023-06-03"),
                            source   = "html",
                            coords   = TRUE,
                            precip_split = TRUE)
head(poznan_html[, c("station_ID", "Date", "TC", "ffkmh", "ddd", "pr6", "pr12", "pr24")])
```

> **Note:** `meteo_ogimet_synop()` is deprecated. Use `meteo_ogimet(source = "synop")` instead.
